---
title: "promotion_assumption_check"
author: "Ke Deng"
date: "2021/4/9"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(readr)
```

```{r}
curr_employee <- read_csv("/Users/krist/Documents/303 final project/data/black-saber-current-employees.csv")
```

### Verify assumptions

- the cases are independently distributed
- the random effect variables are normally distributed

```{r}
qqnorm(curr_employee$employee_id)
qqline(curr_employee$employee_id, col = "orange", lwd=2)
```

From the Quantile-Quantile plot above, the points follow closely around the qq-line in the middle and deviates from it at the two ends. Given this information, the variable "id" can be assumed to be normally distributed, hence matching the model assumption. 

```{r}
cleaned_data <- curr_employee
cleaned_data <- cleaned_data%>%
  mutate(team, team_num =case_when(
    team == "Client services"~1,
    team == "Data"~2,
    team == "Design Legal and financial"~3,
    team == "Marketing and sales"~4,
    team == "Operations"~5,
    team == "People and talent"~6,
    team == "Software"~7  ))

qqnorm(cleaned_data$team_num)
qqline(cleaned_data$team_num, col = "orange", lwd=2)
```
Creating and checking the y-variable:
```{r}
cleaned_data <- curr_employee
#cleaned_data <- cleaned_data%>%
#  group_by(employee_id) %>%
#  count()
cleaned_data <- cleaned_data%>%
  mutate(role_seniority, role_num =case_when(
    role_seniority == "Entry-level"~1,
    role_seniority == "Junior I"~2,
    role_seniority == "Junior II"~3,
    role_seniority == "Senior I"~4,
    role_seniority == "Senior II"~5,
    role_seniority == "Senior III"~6,
    role_seniority == "Manager"~7,
    role_seniority == "Director"~8,
    role_seniority == "Vice president"~9,
    ))

get_promoted=c(0)
for (n in 1:6905){
   if (cleaned_data$employee_id[n] == cleaned_data$employee_id[n+1]){
     if (cleaned_data$role_num[n] == cleaned_data$role_num[n+1]){
       get_promoted[n+1] = 0
     }
     else{
       get_promoted[n+1] = 1
     }
   }
  else{
    get_promoted[n+1]=0
  }
}
cleaned_data$get_promotion = get_promoted
cleaned_data[1,]

table_promotion = table(cleaned_data$get_promotion)
table_promotion

cleaned_data <- cleaned_data%>%
  filter(gender!="Prefer not to say")


model1<- glm(get_promotion~gender+team+leadership_for_level+productivity, data=cleaned_data, family = binomial())
summary(model1)
```


